---
title: "Fars Functions Usage"
author: "KL"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fars Functions Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(utils)
library(readr)
library(dplyr)
library(tidyr)
library(maps)
library(graphics)
```
```{r  global_options}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, message = TRUE,
  warning = FALSE)
utils::globalVariables(c("STATE","MONTH","year"))
```
# Accident data visualization codes and data.

This package provides example data and functions to read in and plot accident data for a given year and state in US.
Raw data files for years 2013, 2014, and 2015 in system.file("extdata", "accident_2013.csv", package = "testi").

## Function to read in data.

This function is given a filename that has both path and name of dataset to download using read_csv function.
It first checks if the data exist, provides an error message if they do not, and downloads the data
if they do in a data.frame (or tibble) format. Not that original Coursera code has here been changed to use 
tibble::as_tibble() instead of dplyr::tbl_df().

Function parameters: filename, a character string giving the path and name of data set to download.

Imported packages are readr, dplyr

Examples: fars_read("accident_2013.csv.bz2")

```{r fars_read}
fars_read <- function(filename) {
        if(!file.exists(filename))
                stop("file '", filename, "' does not exist")
        data <- suppressMessages({
                readr::read_csv(filename, progress = FALSE)
        })
        tibble::as_tibble(data)
}
fars_read("accident_2013.csv.bz2")
```
## Function to crate a filename.

This function returns a filename in C-style string formatting from a character or integer year given by the user and a pre set file name. If the year is provided as a character string, the function will convert it to integer before returning the filename.

Parameters:  year, A character string or an integer provided by user that specifies the year of the file name.

Examples:
 make_filename(2014)
 make_filename("2014")

```{r make_filename}
make_filename <- function(year) {
        year <- as.integer(year)
        sprintf("accident_%d.csv.bz2", year)
}
make_filename(2013)
```
## Function to download multiple files.

This function will download one to several years of data on one go from a list of years given by user.
In case of any error, the function prints warnings of invalid years and returns NULL.

Parameters: years, a vector or list of years given by user that specifies files to download.
Function returns a list of data frames of specified years.

Imported packages: dplyr, tidyr

Examples:
 fars_read_years(c("2013","2014","2015"))
 fars_read_years(c(2013,2014,2015))

```{r fars_read_years}
fars_read_years <- function(years) {
        lapply(years, function(year) {
                file <- make_filename(year)
                tryCatch({
                        dat <- fars_read(file)
                        dplyr::mutate(dat, year = year) %>%
                                dplyr::select(MONTH, year)
                }, error = function(e) {
                        warning("invalid year: ", year)
                        return(NULL)
                })
        })
}
fars_read_years(c("2013","2014"))
```
## Function to summarize multiple years of data.

This function reads in given years of data and returns a summary of n by years and months.

Parameters: years, a vector or list of years to download and summarize.

Imported packages: dplyr, tidyr

Examples:
 fars_summarize_years(c("2013","2014","2015"))
 fars_summarize_years(c(2013,2014,2015))

```{r fars_summarize_years}
fars_summarize_years <- function(years) {
        dat_list <- fars_read_years(years)
        dplyr::bind_rows(dat_list) %>%
                dplyr::group_by(year, MONTH) %>%
                dplyr::summarize(n = n()) %>%
                tidyr::spread(year, n)
}
fars_summarize_years(c(2013,2014))
```
## Function to output a map of road accidents by state.

This function prints a map of road accidents for a state and year given by users. Function prints "invalid state number" if state number does not exist or there are no accidents.

Parameters: state.num, integer or character; year, integer or character

Imported packages: dplyr, maps, graphics

Examples:
 fars_map_state(1,"2014")
 fars_map_state(1,2014)

```{r fars_map_state}
fars_map_state <- function(state.num, year) {
        filename <- make_filename(year)
        data <- fars_read(filename)
        state.num <- as.integer(state.num)

        if(!(state.num %in% unique(data$STATE)))
                stop("invalid STATE number: ", state.num)
        data.sub <- dplyr::filter(data, STATE == state.num)
        if(nrow(data.sub) == 0L) {
                message("no accidents to plot")
                return(invisible(NULL))
        }
        is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
        is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
        with(data.sub, {
                maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
                          xlim = range(LONGITUD, na.rm = TRUE))
                graphics::points(LONGITUD, LATITUDE, pch = 46)
        })
}
fars_map_state(1,2014)
```
